---
title: "Case-II: Lifestyle and Sleeping Quality"
format:
  html:
    toc: true
    toc-location: left  
    toc-depth: 3
    number-sections: true
    smooth-scroll: true
author: Egor Barsegov,

---




# Case study sleep (task 2)

## Part 1



```{r}
old<-par(no.readonly = TRUE)

coljack<- matrix(c("#33ccff","#33ffbb","#33ff33",
                   "#99ff33","#ff9933","#ff3333",
                   "#ff3399","#ff33ff","#7733ff",
                   "#3333ff","#ffff1a"), nrow=11)

```


---

First 6 row of the data 


```{r}
#Reading the file 
sleep<-read.csv("sleep.csv")
head(sleep)
```

---

 Describing the main data set

```{r}
#types of data and sample size 
str(sleep)


#changing tyoes of data to Factor for easier summary 
sleep$occupation <-factor(sleep$occupation)
sleep$sleep_disorder <- factor(sleep$sleep_disorder)


#summary for main data
summary(sleep)



#distribution of age, sleep duration, stress level, daily steps

sapply(sleep[,c("age","sleep_duration","stress_level","daily_steps")], FUN=sd)



#Types of Jobs at a study and length 
unique(sleep$occupation)
length(unique(sleep$occupation))



  
  
  sleepocdu<-aggregate( sleep_duration ~ occupation, data = sleep, FUN = mean) 
  sleepocdu<-cbind(sleepocdu,coljack)
  sleepocdu <- sleepocdu[order(-sleepocdu$sleep_duration),]
  print(sleepocdu)
  
  par(mai = c(0.1, 0.4, 0.1, 0.2),mar = c(11, 4, 2, 2))  
  
  bpsd <- barplot(
    sleepocdu$sleep_duration, col = sleepocdu$coljack, main = "Sleep duration based on Profession", axis.lty = 0, ylim = c(0, 10), ylab = "Sleep (Hours)", names.arg = sleepocdu$occupation, las=2
  )
  


```














---

<br>
<br>
<br>
<br>
<br>
<br>

<br>
<br>

---

## Part 2



<br>
<br>



 



### 2a Relation between Occupation and Stress level, Sleep quality

---

#### Sleep Quality and Occupation


```{r, fig.height=8, fig.width=10}

sleepocqw<-aggregate( sleep_quality ~ occupation, data = sleep, FUN = mean)
sleepocqw<-cbind(sleepocqw,coljack)
sleepocqw <- sleepocqw[order(-sleepocqw$sleep_quality),]
print(sleepocqw[,c("occupation","sleep_quality")])
par(mar = c(10, 4, 4, 2))
barplot(sleepocqw$sleep_quality,
        names.arg = sleepocqw$occupation,
        col = sleepocqw$coljack,
        main = "Sleep quality based on Occupation",
        axis.lty = 0,
        las=2,
        ylim = c(0,10),
        ylab = "Sleep quality (1-10)"
        
        )
        


```

#### Stress Level and Occupation

```{r, fig.height=8, fig.width=10}
sleepocst<-aggregate(stress_level ~ occupation, data = sleep, FUN = mean )
sleepocst<-cbind(sleepocst,coljack)
sleepocst<-sleepocst[order(-sleepocst$stress_level),]
print(sleepocst[,c("occupation","stress_level")])

bpst<-barplot(sleepocst$stress_level, col = sleepocst$coljack,  main = "Occupation and stress level", ylab = "Stress level (1-10)", ylim = c(0,10), )

text(x = bpst, y = sleepocst$stress_level + 0.3, labels = sleepocst$occupation, srt = 90,adj = 0,cex = 1
)
```



### 2b Sleep disorder

```{r}

table(sleep$sleep_disorder)
gender_w_dis<-table(sleep$gender, sleep$sleep_disorder)


par(mfrow=c(1,2),mar= c(2,2,2,4))
pie(gender_w_dis["Male", ],main = "Male")

pie(gender_w_dis["Female", ],main = "Female")

par(old)
``` 


Sleep Apnea is most common overall (excluding no disorder)

Most common in Female is Apnea
Most common in Male is Insomnia


### 2c Combination of Occupation gender Quality 

```{r}
occ_gen_qual <- aggregate(sleep_quality ~ occupation + gender, data = sleep, FUN = mean)
counts <- aggregate(sleep_quality ~ occupation + gender, data = sleep, FUN = length)
names(counts)[3] <- "n"
occ_gen_qual <- merge(occ_gen_qual, counts, by = c("occupation", "gender"))
occ_gen_qual <- occ_gen_qual[order(-occ_gen_qual$sleep_quality), ]
occ_gen_qual



```

```{r}
  sleep_male<-sleep[sleep$gender %in% c("Male"),]
    sleep_female<-sleep[sleep$gender %in% c("Female"),]
```



Lowsest Male sleep quality is for Sales Representatives 


Lowest Female sleep quality is for Scientist

```{r}

sleepocst_Male<-aggregate(stress_level ~ occupation, data = sleep_male, FUN = mean )
sleepocst_Male<-sleepocst_Male[order(-sleepocst_Male$stress_level),]
print(sleepocst_Male)


```
Male Sales Representatives has highest stress



```{r}
sleepocst_Female<-aggregate(stress_level ~ occupation, data = sleep_female, FUN = mean )
sleepocst_Female<-sleepocst_Female[order(-sleepocst_Female$stress_level),]
print(sleepocst_Female)



```
Female Scientist has highest stress


```{r}
print(cor(sleep$sleep_quality,sleep$stress_level))

```
There is a negative correlation between stress level and sleep quality meaning the more stress a person have the worse is there sleep quality. Also basing that there is a similar correlation in separate female and male group i think that there is enough evidence to say there can be a strong correlation. 


---

<br>
<br>
<br>
<br>
<br>
<br>

<br>
<br>

---

## Part 3


<br>
<br>






### a) Statment: 41% of males suffer from insomnia.

```{r}

 sleep_male_insomnia<-sleep_male[sleep_male$sleep_disorder %in% c("Insomnia"),]
print(length(sleep_male_insomnia$gender)/length(sleep_male$gender)*100)

```

Wrong 21.69 percent of males struggle from Insomnia 

### b) Statment: The average daily steps of men who suffer from insomnia is 7065.


```{r}
print(mean(sleep_male_insomnia$daily_steps))
```

Wrong the average daily steps by males with insomnia is 5812

---

<br>
<br>
<br>
<br>
<br>
<br>

<br>
<br>

---

## Part 4 (Box plots)



<br>
<br>



### a) sleep duration and sleepduration 

#### 1.

```{r}
boxplot(sleep_duration ~ sleep_disorder, data =sleep,
        horizontal = TRUE, col = "Lightblue", 
        main= "SLeep duration based on sleep disorders", 
        ylab= "Sleep disorder", xlab= "Sleep duration (hours)")
```

### a) Sleep disorders and Age

#### 2.

```{r}

boxplot(age ~ sleep_disorder, data =sleep,
        horizontal = TRUE, col = "Lightgreen", 
        main= "Sleep disorders based on Age", 
        ylab= "Sleep disorder", xlab= "Age (years)")
```

### b) Computing overlap of confidence intervals 



Looking for intercept at alpha set at 5

```{r}
 sleep_insomnia<-sleep[sleep$sleep_disorder %in% c("Insomnia"),]
  sleep_apnea<-sleep[sleep$sleep_disorder %in% c("Sleep Apnea"),]
  sleep_healthy<-sleep[sleep$sleep_disorder %in% c("None"),]
#sleep intervals for insomnia 
 t.test( sleep_insomnia$sleep_duration,conf.level = 0.95)$conf.int
 
 #sleep intervals for apnea 
 t.test(sleep_apnea$sleep_duration,conf.level = 0.95)$conf.int
 
 #sleep intervals for healthy
 t.test( sleep_healthy$sleep_duration,conf.level = 0.95)$conf.int

```






As it can be seen there is no intercept at alpha = 5 that the result aren't random and have a correlation. 



Now we can try for alpha = 1 

```{r}
#sleep intervals for insomnia 
 t.test( sleep_insomnia$sleep_duration,conf.level = 0.99)$conf.int
 
 #sleep intervals for apnea 
 t.test(sleep_apnea$sleep_duration,conf.level = 0.99)$conf.int

 #sleep intervals for healthy
 t.test( sleep_healthy$sleep_duration,conf.level = 0.99)$conf.int

```
there is a slight intercept between apnea and healthy boundries howeevr only by 0.02 which isnt very significant 

### c) doing part b for age 

Now we repeat but for age 
```{r}
#For healthy 
t.test( sleep_healthy$age,conf.level = 0.95) $conf.int
#For Insomnia 
t.test( sleep_insomnia$age,conf.level = 0.95) $conf.int
#For Apnea 
t.test(sleep_apnea$age,conf.level = 0.95) $conf.int
  
   

```
As it can be seen there is no intercept at alpha = 5 that the result aren't random and have a correlation. 

Now we try for alpha = 1

```{r}
#For healthy 
t.test( sleep_healthy$age,conf.level = 0.99) $conf.int
#For Insomnia 
t.test( sleep_insomnia$age,conf.level = 0.99) $conf.int
#For Apnea 
t.test(sleep_apnea$age,conf.level = 0.99) $conf.int
  
   

```

Even at alpha 1 there is no overlap

This means that, people with no sleep disorder are the youngest on average, people with insomnia are older, and people with sleep apnea are the oldest on average.

Because the confidence intervals do not overlap, this suggests that the age differences between the groups are unlikely to be random.







