---
title: "Case-II: Lifestyle and Sleeping Quality"
format:
  html:
    toc: true
    toc-location: left   # or right
    toc-depth: 3
    number-sections: true
    smooth-scroll: true
author: Egor Barsegov,

---

# Case study sleep (task 2)


## Part 1

```{r}
old<-par(no.readonly = TRUE)

coljack<- matrix(c("#33ccff","#33ffbb","#33ff33","#99ff33","#ff9933","#ff3333","#ff3399","#ff33ff","#7733ff","#3333ff","#ffff1a"), nrow=11)

```



```{r}
#Reading the file 
sleep<-read.csv("sleep.csv")
head(sleep)
```



```{r}
#types of data and sample size 
str(sleep)


#changing tyoes of data to Factor for easier summary 
sleep$occupation <-factor(sleep$occupation)
sleep$sleep_disorder <- factor(sleep$sleep_disorder)


#summary for main data
summary(sleep)



#distribution of age, sleep duration, stress level, daily steps

sapply(sleep[,c("age","sleep_duration","stress_level","daily_steps")], FUN=sd)



#Types of Jobs at a study and length 
unique(sleep$occupation)
length(unique(sleep$occupation))



  
  
  sleepocdu<-aggregate( sleep_duration ~ occupation, data = sleep, FUN = mean) 
  sleepocdu<-cbind(sleepocdu,coljack)
  sleepocdu <- sleepocdu[order(-sleepocdu$sleep_duration),]
  print(sleepocdu)
  
  par(mai = c(0.1, 0.4, 0.1, 0.2),mar = c(11, 4, 2, 2))  
  
  bpsd <- barplot(
    sleepocdu$sleep_duration, col = sleepocdu$coljack, main = "Sleep duration based on Profession", axis.lty = 0, ylim = c(0, 10), ylab = "Sleep (Hours)", names.arg = sleepocdu$occupation, las=2
  )
  


```






## Part 2


### 2a Relation between Occupation and Stress level, Sleep quality

#### Sleep Quality and Occupation

```{r, fig.height=8, fig.width=10}

sleepocqw<-aggregate( sleep_quality ~ occupation, data = sleep, FUN = mean)
sleepocqw<-cbind(sleepocqw,coljack)
sleepocqw <- sleepocqw[order(-sleepocqw$sleep_quality),]
print(sleepocqw[,c("occupation","sleep_quality")])
par(mar = c(10, 4, 4, 2))
barplot(sleepocqw$sleep_quality,
        names.arg = sleepocqw$occupation,
        col = sleepocqw$coljack,
        main = "Sleep quality based on Occupation",
        axis.lty = 0,
        las=2,
        ylim = c(0,10),
        ylab = "Sleep quality (1-10)"
        
        )
        


```
#### Stress Level and Occupation
```{r, fig.height=8, fig.width=10}
sleepocst<-aggregate(stress_level ~ occupation, data = sleep, FUN = mean )
sleepocst<-cbind(sleepocst,coljack)
sleepocst<-sleepocst[order(-sleepocst$stress_level),]
print(sleepocst[,c("occupation","stress_level")])

bpst<-barplot(sleepocst$stress_level, col = sleepocst$coljack,  main = "Occupation and stress level", ylab = "Stress level (1-10)", ylim = c(0,10), )

text(x = bpst, y = sleepocst$stress_level + 0.3, labels = sleepocst$occupation, srt = 90,adj = 0,cex = 1
)
```



### 2b Sleep disorder

```{r}

table(sleep$sleep_disorder)
gender_w_dis<-table(sleep$gender, sleep$sleep_disorder)


par(mfrow=c(1,2),mar= c(2,2,2,4))
pie(gender_w_dis["Male", ],main = "Male")

pie(gender_w_dis["Female", ],main = "Female")

par(old)
``` 

Sleep Apnea is most common overall (excluding no disorder)

Most common in Female is Apnea
Most common in Male is Insomnia


### 2c Combination of Occupation gender Quality 

```{r}
occ_gen_qual <- aggregate(sleep_quality ~ occupation + gender, data = sleep, FUN = mean)
counts <- aggregate(sleep_quality ~ occupation + gender, data = sleep, FUN = length)
names(counts)[3] <- "n"
occ_gen_qual <- merge(occ_gen_qual, counts, by = c("occupation", "gender"))
occ_gen_qual <- occ_gen_qual[order(-occ_gen_qual$sleep_quality), ]
occ_gen_qual



```

```{r}
  sleep_male<-sleep[sleep$gender %in% c("Male"),]
    sleep_female<-sleep[sleep$gender %in% c("Female"),]
```



Lowsest Male sleep quality is for Sales Representatives 


Lowest Female sleep quality is for Scientist

```{r}

sleepocst_Male<-aggregate(stress_level ~ occupation, data = sleep_male, FUN = mean )
sleepocst_Male<-sleepocst_Male[order(-sleepocst_Male$stress_level),]
print(sleepocst_Male)


```
Male Sales Representatives has highest stress



```{r}
sleepocst_Female<-aggregate(stress_level ~ occupation, data = sleep_female, FUN = mean )
sleepocst_Female<-sleepocst_Female[order(-sleepocst_Female$stress_level),]
print(sleepocst_Female)



```
Female Scientist has highest stress


```{r}
print(cor(sleep$sleep_quality,sleep$stress_level))

```
There is a negative correlation between stress level and sleep quality meaning the more stress a person have the worse is there sleep quality. Also basing that there is a similar correlation in separate female and male group i think that there is enough evidence to say there can be a strong correlation. 






## Part 3

```{r}

 sleep_male_insomnia<-sleep_male[sleep_male$sleep_disorder %in% c("Insomnia"),]
print(length(sleep_male_insomnia$gender)/length(sleep_male$gender)*100)

```

Wrong 21.69 percent of males strugle from Insomnoa 


```{r}
print(mean(sleep_male_insomnia$daily_steps))
```

Wring the average daily steps by males with insomnia is 5812

## Part 4

### a) 

#### 1.

```{r}
sleep_insomnia<-sleep[sleep$sleep_disorder %in% c("Insomnia"),]
summary(sleep_insomnia$age)
par(mfrow=c(1,2),mar= c(5,1,5,1))

boxplot(sleep_insomnia$age, horizontal = TRUE, main = "Age of people with Insomnia", col= "lightblue")




sleep_apnea<-sleep[sleep$sleep_disorder %in% c("Sleep Apnea"),]
print(summary(sleep_apnea$age))

boxplot(sleep_apnea$age, horizontal = TRUE,main = "Age of people with Sleep Apnea", col= "lightgreen")

```

### a) 

#### 2.


```{r}
sleep_insomnia<-sleep[sleep$sleep_disorder %in% c("Insomnia"),]
summary(sleep_insomnia$sleep_duration)
par(mfrow=c(1,2),mar= c(5,1,5,1))

boxplot(sleep_insomnia$sleep_duration, horizontal = TRUE, main = "Duration of sleep with Insomnia",col="lightblue")




sleep_apnea<-sleep[sleep$sleep_disorder %in% c("Sleep Apnea"),]
print(summary(sleep_apnea$sleep_duration))

boxplot(sleep_apnea$sleep_duration, horizontal = TRUE,main = "Duration of sleep with Sleep Apnea", col= "lightgreen")

```
